### 2015-06-15 Docs/paper.tex
doc chg

%\documentclass[9pt,twocolumn]{IEEEtran}
%\documentclass[12pt,twoside,draft]{IEEEtran}
%\documentclass[12pt,draft]{IEEEtran}
%\documentclass[11pt,draft]{IEEEtran}
%\documentclass[11pt,final]{IEEEtran}
%\documentclass[11pt,ds,final]{my_IEEEtran}
%\documentclass[11pt,onecolumn,draftcls]{IEEEtran}
%\documentclass[11pt]{IEEEtran}
\documentclass[11pt,onecolumn]{IEEEtran}
%\documentclass[draft]{IEEEtran}
%\documentclass[9pt,twocolumn,technote,twoside]{IEEEtran}

\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}

%\topmargin 0.125in

\usepackage{times}
\usepackage{subfigure}
%\usepackage{algorithmic}
\usepackage{algpseudocode}
\usepackage{algorithm}
%\renewcommand{\algorithmicrequire}{\textbf{Input:}}
%\renewcommand{\algorithmicensure}{\textbf{Output:}}

%\usepackage{epsfig}
%\usepackage{graphics,graphicx,fancyhdr,amsfonts,amsmath,color,epic}
\usepackage[dvips]{graphics,graphicx}
%\usepackage{graphicx,graphics}
\DeclareGraphicsExtensions{.jpg,.pdf,.eps,.png}

\newtheorem {lemma}{Lemma}
\newtheorem {theorem}{Theorem}

\begin{document}

\title{An OpenCL Implementation of Wait-Free Sets}

\author{Sudarshan Shinde \\
        Bangalore, INDIA. \\
Email:sudarshan\_shinde@iitbombay.org}

\maketitle

%\markboth{IEEE Transactions On Automatic Control, Vol. XX, No. Y, Month 1999}
%{Murray and Balemi: Using the style file IEEEtran.sty}

%\noindent EDICS No :DSP-BANK, DSP-MULT, DSP-TFSR.

%\thispagestyle{plain}
%\pagestyle{plain}

\begin{abstract}

\end{abstract}

\begin{IEEEkeywords}
wait-free programming, multithreading, gpgpu, opencl.
\end{IEEEkeywords}

\footnotetext{Copyright (c) 2010 IEEE. Personal use of this material is permitted. However, permission to use this material for any other purposes must be obtained from the IEEE by sending a request to pubs-permissions@ieee.org.}

\pagebreak

\section{Implementation Details}

The algorithms described in the literarure do not describe memory management. In perticular, they do not take into consideration, what happens if a node is physically deleted, while another process it using that node.

OpenCL 2.0 defines Shared Virtual Memory (SVM), that could be used to share data between host and the GPU device, without explicit data transfer. However OpenCL does not support dynamic data allocation on GPU side, and requires that size of SVM be pre-defined. This restricts maximum size of a set.

Since an implementation needs to take care of memory management, in this implementation we allocate a node pool of size $N$ in SVM. Though this restricts set size to be less than $N$, it still highlights many implementation aspects of wait-free set implementation in heterogenous computing setting.

\section{Proof of Correctness}
\begin{algorithm}
  \caption{Deletes all marked nodes beyond {\it startRef}}
  \label{alg:deleteMarkedNode}

  \begin{algorithmic}[1]
    \Require
       \Statex startRef:Reference to the node beyond which all
               marked nodes are to be deleted. 
    \Ensure
       \Statex All marked nodes beyond {\it startRef} are deleted.
       \Statex {\it nextRef} returns next unmarked node beyond {\it startRef}.  
       \Statex Returns {\it true} if all marked nodes are deleted,
               {\it false} if node referenced by {\it startRef} gets deleted.

    \Statex   
    \Function{delMarkedNodes}{startRef, nextRef\&}
      \State doneFlag = false;
      \While{(!doneFlag)}    
        \State [nextRef,startMark] := getUnmarkedRef(startRef);
        \If{(!startMark)}
          \State [nextNextRef,nextMark] := getUnmarkedRef(nextRef);
          \If{(nextMark)}
            \State status = CAS(nodePool[startRef]:next,[nextRef,startMark],[nextNextRef,startMark]);
            \If{(status = true)}
               \State freeNode(nextRef);
            \EndIf
          \Else
            \State
            \Return true;
          \EndIf
        \Else{}
          \State
          \Return false;  
        \EndIf
      \EndWhile        
    \EndFunction
  \end{algorithmic}
\end{algorithm}

\begin{theorem}
  A process running algorithm \ref{alg:deleteMarkedNode}, then it will either return next unmarked node after {\it startRef} in {\it nextRef}, or return {\it false}.
\end{theorem}
\begin{IEEEproof}
  The proof consists of the following parts.
  \begin{enumerate}
    \item The state of the linked list copied to internal variables {\it nextRef} and {\it nnRef} is valid as they are read simultaneously with {\it startMark} and {\it nextMark} respectively, and both these flags are checked.
    \item If this state remains valid, then CAS operation is successful.
    \item If this state is changed by any other process, this change is completely reflected in {\it startRef:next}.  
\end{enumerate} 
\end{IEEEproof}

\begin{algorithm}
  \caption{Check next node for key}
  \label{alg:check}
  \begin{algorithmic}[5]
    \Function{checkKey}{key, prevRef\&, nodePool}
       \State [nextRef,prevMark] := getPtr(prevRef);
       \If {(prevMark)}
         \State
         \Return {\small NODE\_DELETED}.
       \EndIf
       \If {(nextRef = 0)}
         \State
         \Return {\small KEY\_NOT\_DELETED}.
       \EndIf

       \State [nnRef,nextMark] := getPtr(nextRef);
       \If {(nextMark)}
         \State
         \Return {\small NODE\_NEXT}.
       \EndIf

       \State maxVal := simdMax(nodePool,nextRef);
         \If{maxVal $<$ key}
            \State prevRef := nextRef;
            \Return {\small NODE\_NEXT}.
         \Else
           \If{simdAny(val[lid] = key)}
               \State
               \Return {\small KEY\_FOUND};
           \Else
               \State               
               \Return {\small KEY\_NOT\_FOUND};   
           \EndIf   
         \EndIf
    \EndFunction
  \end{algorithmic}
\end{algorithm}

\begin{algorithm}
  \caption{Find a key in the set}
  \label{alg:find}
  \begin{algorithmic}[5]
    \Function{find}{key, prevRef\&, nodePool}
       \State lid := simdGetIndex();
       \State startRef := hashFunction(key);
       \State prevRef  := startRef;
       \State doneFlag := false;
       \While{(!doneFlag)}
          \If{(lid = 1)}
             \State delStatus := delMarkedNode(prevRef,nodePool);
          \EndIf
          \State simdBarrier();
          \State delStatus = simdBroadcast(delStatus,1);
          \If{delStatus = true}
             \State keyStatus := checkKey(key, prevRef, nodePool);
             \If{(keyStatus = NODE\_DELETED)}
                \State prevRef := startRef;
             \ElsIf{(keyStatus = KEY\_FOUND)}
                \State 
                \Return true
             \ElsIf{(keyStatus = KEY\_NOT\_FOUND)}
                \State 
                \Return false
             \EndIf
          \Else
             \State prevRef := startRef;   
          \EndIf
       \EndWhile
    \EndFunction
  \end{algorithmic}
\end{algorithm}

\begin{algorithm}
  \caption{Insert a key in the set}
  \label{alg:insert}
  \begin{algorithmic}[5]  
    \Function{insert}{key, nodePool}
      \State doneFlag := false;
      \While{(!doneFlag)}
        \State findStatus := find(key, prevRef\&, nextRef\&, nodePool);
        \If{(findStaus = true)}
          \State
          \Return false;
        \EndIf
        \State minIndex := getMinEmptySlot(nextRef,nodePool);
        \If{minIndex = 0} \Comment No empty slot.
          \State insertStatus := insertNodeAndKey(key, prevRef\&,nextRef\&,nodePool);
        \Else
          \State insertStatus := insertKey(key, minIndex, nextRef\&,nodePool);
        \EndIf
        \If{insertStatus = true}
          \State
          \Return true;
        \EndIf
      \EndWhile
    \EndFunction
  \end{algorithmic}
\end{algorithm}

\begin{algorithm}
  \caption{Find empty slot with minimum index}
  \label{alg:minindex}
  \begin{algorithmic}[5]  
    \Function{getMinEmptySlot}{nextRef, nodePool}
      \State val := nodePool[nextRef]:V[lid];
      \State indicator := (val = EMPTY\_SLOT)?lid:0;
      \If {(simdALL(indicator = 0))}
        \State minIndex := 0;
      \Else
        \State minIndex := simdMin(indicator);
      \EndIf
      \Return minIndex;        
    \EndFunction
  \end{algorithmic}
\end{algorithm}

\begin{algorithm}
  \caption{Insert a node and a key}
  \label{alg:insnodekey}
  \begin{algorithmic}[5]  
    \Function{insertNodeAndKey}{key, prevRef\&, nextRef\&, nodePool}
      \State newNRef := Alloc(nodePool);
      \State newNNRef := Alloc(nodePool);
      \State newNRef:next := [newNNRef,0];
      \If{(nodePool[nextRef]:V[lid] < key)}
        \State nodePool[newNRef]:V[lid] := nodePool[nextRef]:V[lid];
        \State nodePool[newNNRef]:V[lid] := EMPTY\_SLOT;
      \Else
        \State nodePool[newNNRef]:V[lid] := nodePool[nextRef]:V[lid];
        \State nodePool[newNPRef]:V[lid] := EMPTY\_SLOT;
      \EndIf
      \State minIndex := getMinEmptySlot(newNRef,nodePool);
      \State nodePool[newNRef]:V[minIndex] := key;
      \State nodePool[newNNPtr]:next := [nextRef,0];
      \State status := CAS(nodePool[prevRef]:next,[nextRef,0],newNPtr);
      \Return status;
    \EndFunction
  \end{algorithmic}
\end{algorithm}

\begin{algorithm}
  \caption{Insert a key in a node}
  \label{alg:inskey}
  \begin{algorithmic}[5]  
    \Function{insertKey}{key, minIndex, nextRef\&, nodePool}
      \If{lid = 1}
        \State status := CAS(nodePool[nextRef]:V[minIndex],EMPTY\_SLOT,key);
      \EndIf
      \State simdWait();
      \State status = simdBroadcast(status, 1);
      \Return status;
    \EndFunction
  \end{algorithmic}
\end{algorithm}

\begin{algorithm}
  \caption{Delete a key from the set}
  \label{alg:delkey}
  \begin{algorithmic}[5]  
    \Function{remove}{key, nodePool}
      \State doneFlag := false;
      \While{(!doneFlag)}
        \State findStatus := find(key, prevRef\&, nextRef\&, nodePool);
        \If{(findStaus = false)}
          \State
          \Return false;
        \EndIf
        \State val   := nodePool[nextRef]:V[lid];
        \If{(val = key)}
          \State status := CAS(nodePool[nextRef]:V[lid],key,EMPTY\_SLOT);
        \EndIf
        \State simdWait();
        \State status := simdBrodacast(status);
        \If{(status = true)}
          \State status := deleteNode(prevRef\&, nextRef\&, nodePool);
          \Return status;
        \Else
          \State
          \Return false;
        \EndIf
      \EndWhile
    \EndFunction
  \end{algorithmic}
\end{algorithm}

\begin{algorithm}
  \caption{Delete a node}
  \label{alg:delnode}
  \begin{algorithmic}[5]  
    \Function{deleteNode}{prevRef\&, nextRef\&, nodePool}
         \State val := nodePool[nextRef]:V[lid];
         \State status := simdAll(val,EMPTY\_SLOT);
         \If{(status = true)}
           \State [nnRef,nextMark] := getPtr(nextRef);
           \State newMark := 1;
           \If{(lid = 1)}
              \State status := CAS(nodePool[nextRef]:next,[nnRef,nextMark],[nnRef,newMark]);
              \If{(status = true)}
                \State status = delMarkedNodes(prevRef,NodePool); 
              \EndIf
           \EndIf
           \State simdWait();
           \State status = simdBroadcast(status,1);
           \State
           \Return status;
         \EndIf
    \EndFunction
  \end{algorithmic}
\end{algorithm}

%\Bibliographystyle{ieeetr}
%\include{references.bbl}

%%%%%\bibliography{bib-file}  % commented if *.bbl file included, as seen below

\end{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%  End of IEEEsample.tex  %%%%%%%%%%%%%%%%%%%%%%%%%%%