### 2015-06-19 Docs/paper.tex
doc changed

%\documentclass[9pt,twocolumn]{IEEEtran}
%\documentclass[12pt,twoside,draft]{IEEEtran}
%\documentclass[12pt,draft]{IEEEtran}
%\documentclass[11pt,draft]{IEEEtran}
%\documentclass[11pt,final]{IEEEtran}
%\documentclass[11pt,ds,final]{my_IEEEtran}
%\documentclass[11pt,onecolumn,draftcls]{IEEEtran}
%\documentclass[11pt]{IEEEtran}
\documentclass[11pt,onecolumn]{IEEEtran}
%\documentclass[draft]{IEEEtran}
%\documentclass[9pt,twocolumn,technote,twoside]{IEEEtran}

\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}

%\topmargin 0.125in

\usepackage{times}
\usepackage{subfigure}
%\usepackage{algorithmic}
\usepackage{algpseudocode}
\usepackage{algorithm}
%\renewcommand{\algorithmicrequire}{\textbf{Input:}}
%\renewcommand{\algorithmicensure}{\textbf{Output:}}

%\usepackage{epsfig}
%\usepackage{graphics,graphicx,fancyhdr,amsfonts,amsmath,color,epic}
\usepackage[dvips]{graphics,graphicx}
%\usepackage{graphicx,graphics}
\DeclareGraphicsExtensions{.jpg,.pdf,.eps,.png}

\newtheorem {lemma}{Lemma}
\newtheorem {theorem}{Theorem}

\begin{document}

\title{An OpenCL Implementation of Wait-Free Sets}

\author{Sudarshan Shinde \\
        Bangalore, INDIA. \\
Email:sudarshan\_shinde@iitbombay.org}

\maketitle

%\markboth{IEEE Transactions On Automatic Control, Vol. XX, No. Y, Month 1999}
%{Murray and Balemi: Using the style file IEEEtran.sty}

%\noindent EDICS No :DSP-BANK, DSP-MULT, DSP-TFSR.

%\thispagestyle{plain}
%\pagestyle{plain}

\begin{abstract}

\end{abstract}

\begin{IEEEkeywords}
wait-free programming, multithreading, gpgpu, opencl.
\end{IEEEkeywords}

\footnotetext{Copyright (c) 2010 IEEE. Personal use of this material is permitted. However, permission to use this material for any other purposes must be obtained from the IEEE by sending a request to pubs-permissions@ieee.org.}

\pagebreak

\section{Implementation Details}

The algorithms described in the literarure do not describe memory management. In perticular, they do not take into consideration, what happens if a node is physically deleted, while another process it using that node.

OpenCL 2.0 defines Shared Virtual Memory (SVM), that could be used to share data between host and the GPU device, without explicit data transfer. However OpenCL does not support dynamic data allocation on GPU side, and requires that size of SVM be pre-defined. This restricts maximum size of a set.

Since an implementation needs to take care of memory management, in this implementation we allocate a node pool of size $N$ in SVM. Though this restricts set size to be less than $N$, it still highlights many implementation aspects of wait-free set implementation in heterogenous computing setting.

\section{Proof of Correctness}
\begin{algorithm}
  \caption{Deletes all marked nodes that follow a given node}
  \label{alg:deleteMarkedNode}

  \begin{algorithmic}[1]
    \Require
        \Statex startRef:Reference to the starting node. Marked nodes
                that follow this are to be deleted. 
    \Ensure
       \Statex All marked nodes following {\it startRef} are deleted.
       \Statex Returns {\it true} if all marked nodes are deleted,
               {\it false} if node referenced by {\it startRef} gets deleted.

    \Statex   
    \Function{delMarkedNodes}{startRef}
      \State doneFlag = false;
      \While{(!doneFlag)}    
        \State [nextRef,startMark] := getUnmarkedRef(startRef);
        \If{(!startMark)}
          \State [nextNextRef,nextMark] := getUnmarkedRef(nextRef);
          \If{(nextMark)}
            \State status = CAS(startRef:next,[nextRef,startMark],[nextNextRef,startMark]);
            \If{(status = true)}
               \State freeNode(nextRef);
            \EndIf
          \Else
            \State
            \Return true;
          \EndIf
        \Else{}
          \State
          \Return false;  
        \EndIf
      \EndWhile        
    \EndFunction
  \end{algorithmic}
\end{algorithm}

\begin{theorem}
  A process running algorithm \ref{alg:deleteMarkedNode}, then it will either return next unmarked node after {\it startRef} in {\it nextRef}, or return {\it false}.
\end{theorem}
\begin{IEEEproof}
  The proof consists of the following parts.
  \begin{enumerate}
    \item The state of the linked list copied to internal variables {\it nextRef} and {\it nnRef} is valid as they are read simultaneously with {\it startMark} and {\it nextMark} respectively, and both these flags are checked.
    \item If this state remains valid, then CAS operation is successful.
    \item If this state is changed by any other process, this change is completely reflected in {\it startRef:next}.  
\end{enumerate} 
\end{IEEEproof}

\begin{algorithm}
  \caption{Check next node for key}
  \label{alg:check}
  \begin{algorithmic}[1]
    \Require
       \Statex {\it key} to be checked.
       \Statex {\it prevRef}, reference to node next to which the key is to
               be checked.
    \Ensure
       \Statex Nodes referenced by {\it prevRef} and {\it prevRef:next} are
               unmarked. If they are marked, returns NODE\_DELETED and
               NODE\_NEXT respectively.
       \Statex If all values in {\it prevRef:next} $<$ {\it key}, moves
               {\it prevRef} to {\it prevRef:next}.          
       \Statex If values in {\it prevRef:next} $<$ {\it key}, checks for the key.
               Returns KEY\_FOUND or KEY\_NOT\_FOUND respectively.

    \Statex   
    \Function{checkKey}{key, prevRef\&}
       \State [nextRef,prevMark] := getUnmarkedPtr(prevRef);
       \If {(prevMark)}
         \State
         \Return {\small NODE\_DELETED}.
       \EndIf
       \If {(nextRef = 0)}
         \State
         \Return {\small KEY\_NOT\_FOUND}.
       \EndIf

       \State [nnRef,nextMark] := getUnmarkedPtr(nextRef);
       \If {(nextMark)}
         \State
         \Return {\small NODE\_NEXT}.
       \EndIf

       \State maxVal := simdMAX(nextRef:V[lid]);
       \If{maxVal $<$ key}
           \State prevRef := nextRef;
           \Return {\small NODE\_NEXT}.
       \Else
           \If{simdANY(nextRef:V[lid] = key)}
               \State
               \Return {\small KEY\_FOUND};
           \Else
               \State               
               \Return {\small KEY\_NOT\_FOUND};   
           \EndIf   
         \EndIf
    \EndFunction
  \end{algorithmic}
\end{algorithm}

\begin{algorithm}
  \caption{Find a key in the set}
  \label{alg:find}
  \begin{algorithmic}[1]
    \Require
       \Statex {\it key} to be searched.
    \Ensure
       \Statex {\it prevRef}, such that next node either has key or key could
               be inserted to it.
       \Statex Returns {\it true} if key is found else returns {\it false}.

    \Function{find}{key, prevRef\&}
       \State lid := simdINDEX();
       \State startRef := hashFunction(key);
       \State prevRef  := startRef;
       \State doneFlag := false;
       \While{(!doneFlag)}
          \If{(lid = 1)}
             \State delStatus := delMarkedNode(prevRef);
          \EndIf
          \State simdBARRIER();
          \State delStatus = simdBROADCAST(delStatus,1);
          \If{delStatus = true}
             \State keyStatus := checkKey(key, prevRef);
             \If{(keyStatus = NODE\_DELETED)}
                \State prevRef := startRef;
             \ElsIf{(keyStatus = KEY\_FOUND)}
                \State 
                \Return true
             \ElsIf{(keyStatus = KEY\_NOT\_FOUND)}
                \State 
                \Return false
             \EndIf
          \Else
             \State prevRef := startRef;   
          \EndIf
       \EndWhile
    \EndFunction
  \end{algorithmic}
\end{algorithm}

\begin{algorithm}
  \caption{Insert a key in the set}
  \label{alg:insert}
  \begin{algorithmic}[1]
    \Require
       \Statex {\it key} to be searched.
    \Ensure
       \Statex Returns {\it false} if key is found else inserts the {\it key}.

    \Statex
    \Function{insert}{key}
      \State doneFlag := false;
      \While{(!doneFlag)}
        \State findStatus := find(key, prevRef);
        \If{(findStaus = true)}
          \State
          \Return false;
        \EndIf
        \State [nextRef, prevMark] = getUnmarkedRef(prevRef);
        \If{(prevMark = 0)}
          \State insertStatus = false;
          \State minIndex := getMinEmptySlot(nextRef,nodePool);
          \If{(minIndex = 0)} \Comment No empty slot.
            \State insertStatus := insertNodeAndKey(key, prevRef,nextRef);
          \ElsIf{(minIndex > 0)}
            \State insertStatus := insertKey(key, minIndex, prevRef, nextRef);
          \EndIf
          \If{insertStatus = true}
            \State
            \Return true;
          \EndIf
        \EndIf
      \EndWhile
    \EndFunction
  \end{algorithmic}
\end{algorithm}

\begin{algorithm}
  \caption{Find empty slot with minimum index}
  \label{alg:minindex}
  \begin{algorithmic}[1]
    \Require
      \Statex {\it nextRef} in which minumum index to empty slot is to be
              found.
    \Ensure
      \Statex Returns {\it -1} if {\it nextRef} is deleted, {\it 0} if no
              empty slot is found, or minimum index to empty slot.

    \Function{getMinEmptySlot}{nextRef}
      \State val := nextRef:V[lid];
      \State indicator := (val = EMPTY\_SLOT)?lid:0;
      \If {(simdALL(indicator = 0))}
        \State minIndex := 0;
      \Else
        \State minIndex := simdMIN(indicator);
      \EndIf
      \State [nnRef,nextMark] := getUnmarkedPtr(nextRef);
      \If{(nextMark = 0)}
        \State
        \Return minIndex;
      \Else
        \State
        \Return -1;
      \EndIf  
    \EndFunction
  \end{algorithmic}
\end{algorithm}

\begin{algorithm}
  \caption{Insert a node and a key}
  \label{alg:insnodekey}
  \begin{algorithmic}[1]
    \Require
      \Statex {\it key} to be inserted.
      \Statex {\it prefRef} and {\it nextRef} give window in which new node
              and key are to be inserted.
    \Ensure
       \Statex Returns {\it true} if successful.

    \Function{insertNodeAndKey}{key, prevRef, nextRef}
      \If{(lid = 1)} 
        \State newNRef := newNode();
        \State newNNRef := newNode();
      \EndIf
      \State simdBARRIER();
      \State newNRef  := simdBROADCAST(newNRef,1);
      \State newNNRef := simdBROADCAST(newNRef,1);

      \State newNRef:next := [newNNRef,0];
      \If{(nodePool[nextRef]:V[lid] < key)}
        \State nodePool[newNRef]:V[lid] := nodePool[nextRef]:V[lid];
        \State nodePool[newNNRef]:V[lid] := EMPTY\_SLOT;
      \Else
        \State nodePool[newNNRef]:V[lid] := nodePool[nextRef]:V[lid];
        \State nodePool[newNPRef]:V[lid] := EMPTY\_SLOT;
      \EndIf
      \State minIndex := getMinEmptySlot(newNRef,nodePool);
      \State nodePool[newNRef]:V[minIndex] := key;
      \State nodePool[newNNPtr]:next := [nextRef,0];
      \State status := CAS(nodePool[prevRef]:next,[nextRef,0],newNPtr);
      \Return status;
    \EndFunction
  \end{algorithmic}
\end{algorithm}

\begin{algorithm}
  \caption{Insert a key in a node}
  \label{alg:inskey}
  \begin{algorithmic}[5]  
    \Function{insertKey}{key, minIndex, nextRef\&, nodePool}
      \If{lid = 1}
        \State status := CAS(nodePool[nextRef]:V[minIndex],EMPTY\_SLOT,key);
      \EndIf
      \State simdWait();
      \State status = simdBroadcast(status, 1);
      \Return status;
    \EndFunction
  \end{algorithmic}
\end{algorithm}

\begin{algorithm}
  \caption{Delete a key from the set}
  \label{alg:delkey}
  \begin{algorithmic}[5]  
    \Function{remove}{key, nodePool}
      \State doneFlag := false;
      \While{(!doneFlag)}
        \State findStatus := find(key, prevRef\&, nextRef\&, nodePool);
        \If{(findStaus = false)}
          \State
          \Return false;
        \EndIf
        \State val   := nodePool[nextRef]:V[lid];
        \If{(val = key)}
          \State status := CAS(nodePool[nextRef]:V[lid],key,EMPTY\_SLOT);
        \EndIf
        \State simdWait();
        \State status := simdBrodacast(status);
        \If{(status = true)}
          \State status := deleteNode(prevRef\&, nextRef\&, nodePool);
          \Return status;
        \Else
          \State
          \Return false;
        \EndIf
      \EndWhile
    \EndFunction
  \end{algorithmic}
\end{algorithm}

\begin{algorithm}
  \caption{Delete a node}
  \label{alg:delnode}
  \begin{algorithmic}[5]  
    \Function{deleteNode}{prevRef\&, nextRef\&, nodePool}
         \State val := nodePool[nextRef]:V[lid];
         \State status := simdAll(val,EMPTY\_SLOT);
         \If{(status = true)}
           \State [nnRef,nextMark] := getPtr(nextRef);
           \State newMark := 1;
           \If{(lid = 1)}
              \State status := CAS(nodePool[nextRef]:next,[nnRef,nextMark],[nnRef,newMark]);
              \If{(status = true)}
                \State status = delMarkedNodes(prevRef,NodePool); 
              \EndIf
           \EndIf
           \State simdWait();
           \State status = simdBroadcast(status,1);
           \State
           \Return status;
         \EndIf
    \EndFunction
  \end{algorithmic}
\end{algorithm}

%\Bibliographystyle{ieeetr}
%\include{references.bbl}

%%%%%\bibliography{bib-file}  % commented if *.bbl file included, as seen below

\end{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%  End of IEEEsample.tex  %%%%%%%%%%%%%%%%%%%%%%%%%%%